\chapter*{Abstract}

Cloud Computing is a subject that has recently become very discussed in the academic community. FEUP - Faculty of Engineering from the Oporto University - is no exception. Its informatic center CICA - Professor Correia de Araújo Informatic Center - is currently developing a private Cloud project. This project is split into two different projects and this document covers the front-end of the system.

This projects objectives consist in developing a web portal capable of creating and managing virtual high performance computing environments over private cloud computing infrastructures.

In this document, the state of the art of Cloud Computing related subjects is discussed, and the other half of this project (developed by another student as his Master thesis) is analyzed in various sections.

\chapter*{Resumo}

A computação em nuvem é um tema que recentemente tem vindo a ser muito debatido na comunidade académica. A FEUP - Faculdade de Engenharia da Universidade do Porto - não é excepção. O seu centro de informática CICA - Centro de Informática Prof. Correia de Araújo - encontra-se neste momento a desenvolver um projecto de computação em nuvem privada. O projecto está subdivido em dois projectos e este documento foca-se no front-end do sistema.

Este projecto consiste em desenvolver um portal web capaz de criar e gerir ambientes de computação de elevado desempenho virtuais sobre infraestruturas de computação em nuvem privadas.

Neste documento é exposto o estado da arte da computação em nuvem e temas relacionados, sendo que a outra parte do projecto é analisada em diversas secções do documento.

